---
title: "Day Twenty-Five: APIs"
author: Lindsay Poirier
subtitle: "SDS 192: Introduction to Data Science"
author: |
  Lindsay Poirier<br/>
  <span style = 'font-size: 70%;'>
  [Statistical & Data Sciences](http://www.smith.edu/sds), Smith College<br/>
  </span>
date: |
  Spring 2022<br/>
output: pdf_document
---

```{r}
library(readr)
```

# Create an API call to:

* Access all NYC 311 requests from the Bronx Borough 
* Limited to 100 entries
* As a CSV

> You should refer to the API documentation, our lecture slides, and review the data table [here](https://data.cityofnewyork.us/Social-Services/311-Service-Requests-from-2010-to-Present/erm2-nwe9) to construct the URL. 

```{r}
nyc_311_bronx <- read_csv()
```


# Create an API call to:

* Access all NYC 311 requests created March 1 of last year 
* Limited to 100 entries
* As a CSV

> You should refer to the API documentation, our lecture slides, and review the data table [here](https://data.cityofnewyork.us/Social-Services/311-Service-Requests-from-2010-to-Present/erm2-nwe9) to construct the URL. 

```{r}
nyc_311_mar_1 <- read_csv()
```

# Create an API call to:

* Access all NYC 311 requests of complaint type "Noise - Street/Sidewalk" and descriptor "Loud Music/Party"
* Limited to 100 entries
* As a CSV

> You should refer to the API documentation, our lecture slides, and review the data table [here](https://data.cityofnewyork.us/Social-Services/311-Service-Requests-from-2010-to-Present/erm2-nwe9) to construct the URL. 

> You can reference this link for URL encodings [here](https://www.w3schools.com/tags/ref_urlencode.asp). 

```{r}
# Code here!
nyc_311_loud_music <-
  read_csv()
```

