---
title: "Day Ten: Frequency"
subtitle: "SDS 192: Introduction to Data Science"
author: |
  Lindsay Poirier<br/>
  <span style = 'font-size: 70%;'>
  [Statistical & Data Sciences](http://www.smith.edu/sds), Smith College<br/>
  </span>
date: |
  Spring 2022<br/>
output: pdf_document
---

# Step 1: Install and load packages

In your console install the Spotify R package:

`install.packages("spotifyr")`

Then load the packages for this lab by running the code below. 

```{r}
#Load packates
library(tidyverse)
library(spotifyr)
```

# Step 2: Create an Spotify Developer Account

1. Log-in to Spotify.com or create an account
2. Agree to the Terms and Conditions
3. Create an app named "SDS 192 Class Project". You can indicate that this is a "Project for SDS 192 class"
4. Click "Show Client Secret"
5. Copy client id and secret below, and then run the code chunk.

```{r}
id <- 'COPY_ID_HERE'
secret <- 'COPY_SECRET_HERE'
Sys.setenv(SPOTIFY_CLIENT_ID = id)
Sys.setenv(SPOTIFY_CLIENT_SECRET = secret)
access_token <- get_spotify_access_token()
```

# Step 3: Get song features for your favorite artist

Replace the text FILL FAVORITE ARTIST NAME HERE below with the name of your favorite artist and then run the code chunk. 

```{r}
artist <- get_artist_audio_features(artist = "FILL FAVORITE ARTIST NAME HERE") %>%
  select(-c(album_images, artists, available_markets))
```

# Step 4: Create a histogram visualizing the distribution of values in a song feature of your choice

Be sure to add labels to your plot. 

```{r}
# Fill plot here. 
```

# Step 5: Facet your previous plot by album name

Be sure to add labels to your plot. 

```{r}
# Fill plot here. 
```

Turn to a neighbor and discuss the plots. 
* What do the distributions indicate?
* What do you learn when comparing distributions across albums?

# Step 6: Create a stacked, dodged, or filled barplot visualizing the frequency of songs in each album and key for this artist.

Be sure to add labels to your plot. 

```{r}
# Fill plot here. 
```

Turn to a neighbor and discuss the plots. 
* What do the frequencies indicate?
* What do you learn when comparing frequencies across albums?
