---
title: "Day Eleven: Boxplots"
subtitle: "SDS 192: Introduction to Data Science"
author: |
  Lindsay Poirier<br/>
  <span style = 'font-size: 70%;'>
  [Statistical & Data Sciences](http://www.smith.edu/sds), Smith College<br/>
  </span>
date: |
  Spring 2022<br/>
output: pdf_document
---

# Step 1: Load packages

```{r}
#Load packates
library(tidyverse)
library(spotifyr)
```

# Step 2: Create an Spotify Developer Account

Copy client id and secret from your previous lab into the chunk below, and then run the code chunk.

```{r}
id <- 'COPY_ID_HERE'
secret <- 'COPY_SECRET_HERE'
Sys.setenv(SPOTIFY_CLIENT_ID = id)
Sys.setenv(SPOTIFY_CLIENT_SECRET = secret)
access_token <- get_spotify_access_token()
```

# Step 3: Get song features for your favorite artist

Replace the text FILL FAVORITE ARTIST NAME HERE below with the name of your favorite artist and then run the code chunk. 

```{r}
artist <- get_artist_audio_features(artist = "FILL FAVORITE ARTIST NAME HERE") %>%
  select(-c(album_images, artists, available_markets))
```

# Step 4: Create grouped boxplots visualizing the distribution of values in a song feature of your choice, grouped by album name

> Refer to the [`ggplot()` cheatsheet](https://github.com/rstudio/cheatsheets/blob/main/data-visualization-2.1.pdf) to find the geom function you need to create this plot. 

Be sure to add labels to your plot. 

```{r}
# Fill plot here. 
```

# Step 5: Interpret results

Turn to a neighbor and discuss the plots. 
* Check for outliers.
* Compare medians. 
* Compare the ranges.
* Compare the IQRs.
* Compare the symmetry. 

> Call us over if you are having trouble interpreting!


