---
title: "Day Thirty-Three: How to Lie with Maps"
subtitle: "SDS 192: Introduction to Data Science"
author: |
  Lindsay Poirier<br/>
  <span style = 'font-size: 70%;'>
  [Statistical & Data Sciences](http://www.smith.edu/sds), Smith College<br/>
  </span>
date: |
  Spring 2022<br/>
output: 
  html_document:
    toc: true
    toc_float: true
    theme: lumen
---

## Load Packages

```{r setup,	message = FALSE, warning = FALSE, include = FALSE}
library(readr)
library(leaflet)
library(RColorBrewer)
library(tidyverse)
library(sf)
```

```{r read-voting, include=FALSE}
# Import Voting Precinct Data
voting_2020 <- st_read("datasets/ct_2020.shp") %>%
  st_transform(4326) %>%
  mutate(TOTAL = rowSums(across(starts_with("G20")))) %>%
  mutate(PER_DEM = G20PREDBID / TOTAL * 100) %>%
  mutate(PER_DEM_LEAD = (G20PREDBID - G20PRERTRU) / TOTAL * 100)

voting_2016 <- st_read("datasets/ct_2016.shp") %>%
  st_transform(4326)
```

# Standardization

Below we are going to create two leaflet maps. The first should be colored by the total votes received by Joseph Biden, and the second should be the percentage of votes received by Joseph Biden. 

> Discuss with a neighbor how the maps are different and what each tells us. Is one more honest than another?

```{r dem-map, echo=FALSE}
pal_numeric <- colorNumeric(palette = "Blues", 
                    domain = voting_2020$G20PREDBID)

leaflet() %>%
  setView(lat = 41.6, lng = -71.5, zoom = 8) %>% #Set the View to your state
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(data = voting_2020,
              fillColor = ~pal_numeric(G20PREDBID),
              stroke = FALSE,
              fillOpacity = 0.5) %>%
  addLegend(data = voting_2020,
            title = "Votes for Biden, 2020", 
            pal = pal_numeric, 
            values = ~G20PREDBID)
  
```

```{r per-dem-map, echo=FALSE}
pal_numeric <- colorNumeric(palette = "Blues", 
                    domain = voting_2020$PER_DEM)

leaflet() %>%
  setView(lat = 41.6, lng = -71.5, zoom = 8) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(data = voting_2020,
              fillColor = ~pal_numeric(PER_DEM),
              stroke = FALSE,
              fillOpacity = 0.5) %>%
  addLegend(data = voting_2020,
            title = "% Votes for Biden, 2020", 
            pal = pal_numeric, 
            values = ~PER_DEM)
  
```

# Classification, Part 1

Below we are going to create two leaflet maps. Both will use `colorBin()` to illustrate the percentage of Biden's lead over Trump. However, the first will have 2 bins, and the second will have 8 bins. The domain will be -100:100 (i.e. a -100% lead to a 100% lead). 

> Discuss with a neighbor how the maps are different and what each tells us. Is one more honest than another?

```{r bin-map2, echo=FALSE}
pal_bin <- colorBin(palette = "RdYlBu", 
                    domain = c(-100:100),
                    bins = 2,
                    pretty = FALSE)

leaflet() %>%
  setView(lat = 41.6, lng = -71.5, zoom = 8) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(data = voting_2020,
              fillColor = ~pal_bin(PER_DEM_LEAD),
              stroke = FALSE,
              fillOpacity = 0.5) %>%
  addLegend(data = voting_2020,
            title = "% Biden Lead Over Trump, 2020", 
            pal = pal_bin, 
            values = ~PER_DEM_LEAD)
  
```

```{r bin-map8, echo=FALSE}
pal_bin <- colorBin(palette = "RdYlBu", 
                    domain = c(-100:100),
                    bins = 8,
                    pretty = FALSE)

leaflet() %>%
  setView(lat = 41.6, lng = -71.5, zoom = 8) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(data = voting_2020,
              fillColor = ~pal_bin(PER_DEM_LEAD),
              stroke = FALSE,
              fillOpacity = 0.5) %>%
  addLegend(data = voting_2020,
            title = "% Biden Lead Over Trump, 2020", 
            pal = pal_bin, 
            values = ~PER_DEM_LEAD)
  
```

# Classification, Part 2

Below we are going to create two leaflet maps. The first should be colored by the percentage of Biden's lead over Trump using the `colorBin()` function, and the second should colored by the percentage of Biden's lead over Trump using the `colorQuantile()` function. Both should have 8 bins; however, the domains will be slightly different. For binning, the domain will be -100:100 (i.e. a -100% lead to a 100% lead). For quantiles, the domain will be all of the values in `voting_2020$PER_DEM_LEAD`. Remember:

* `colorBin()` classifies numeric values into equally spaced intervals
* `colorQuantile()` classifies an equal number of numeric values into a specified number of bins

> Discuss with a neighbor how the maps are different and what each tells us. Is one more honest than another?

```{r bin-map, echo=FALSE}
pal_bin <- colorBin(palette = "RdYlBu", 
                    domain = c(-100:100),
                    bins = 8,
                    pretty = FALSE)

leaflet() %>%
  setView(lat = 41.6, lng = -71.5, zoom = 8) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(data = voting_2020,
              fillColor = ~pal_bin(PER_DEM_LEAD),
              stroke = FALSE,
              fillOpacity = 0.5) %>%
  addLegend(data = voting_2020,
            title = "% Biden Lead Over Trump, 2020", 
            pal = pal_bin, 
            values = ~PER_DEM_LEAD)
  
```

```{r quantile-map, echo=FALSE}
pal_quantile <- colorQuantile(palette = "RdYlBu", 
                    domain = voting_2020$PER_DEM_LEAD,
                    n = 8)

leaflet() %>%
  setView(lat = 41.6, lng = -71.5, zoom = 8) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(data = voting_2020,
              fillColor = ~pal_quantile(PER_DEM_LEAD),
              stroke = FALSE,
              fillOpacity = 0.5) %>%
  addLegend(data = voting_2020,
            title = "% Biden Lead Over Trump, 2020", 
            pal = pal_quantile, 
            values = ~PER_DEM_LEAD)
  
```

# Aggregation

Below we are going to create two leaflet maps. For the first you should copy the "% Biden Lead Over Trump, 2020" you created above using `colorBin()`. For the second, we are going to create the same plot aggregated by counties instead of voting precincts. 

> Discuss with a neighbor how the maps are different and what each tells us. Is one more honest than another?

```{r bin-map-repeat, echo=FALSE}
pal_bin <- colorBin(palette = "RdYlBu", 
                    domain = c(-100:100),
                    bins = 8,
                    pretty = FALSE)

leaflet() %>%
  setView(lat = 41.6, lng = -71.5, zoom = 8) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(data = voting_2020,
              fillColor = ~pal_bin(PER_DEM_LEAD),
              stroke = FALSE,
              fillOpacity = 0.5) %>%
  addLegend(data = voting_2020,
            title = "% Biden Lead Over Trump, 2020", 
            pal = pal_bin, 
            values = ~PER_DEM_LEAD)
```


```{r county-map, echo=FALSE, message=FALSE, warning=FALSE}
sf_use_s2(FALSE) #Helps deal with invalid geo problem when aggregating geometry

voting_2020_county <-
  voting_2020 %>%
  group_by(COUNTYFP20) %>%
  summarize_at(vars(starts_with("G20"), TOTAL), sum, na.rm = TRUE) %>%
  ungroup() %>%
  mutate(PER_DEM = G20PREDBID / TOTAL * 100) %>%
  mutate(PER_DEM_LEAD = (G20PREDBID - G20PRERTRU) / TOTAL * 100) %>%
  st_as_sf()
  
pal_bin <- colorBin(palette = "RdYlBu", 
                    domain = c(-100:100),
                    bins = 8, 
                    pretty = FALSE)

leaflet() %>%
  setView(lat = 41.6, lng = -71.5, zoom = 8) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(data = voting_2020_county,
              fillColor = ~pal_bin(PER_DEM_LEAD),
              stroke = FALSE,
              fillOpacity = 0.5) %>%
  addLegend(data = voting_2020_county,
            title = "% Biden Lead Over Trump, 2020", 
            pal = pal_bin, 
            values = ~PER_DEM_LEAD)
  
```


# Reprecincting

Below you are going to create a map with two layers - the first will outline 2016 precinct data, and the second will outline 2020 precinct data. 

> Zoom in on different cities. Do you notice any precinct lines that have been redrawn? Searching online, can you find any information about why these precincts have been redrawn (e.g. search "precinct changes in [CITY NAME]")? What effect do these changes have on how the data gets aggregated?

```{r precinct-map, echo=FALSE, message=FALSE, warning=FALSE}
leaflet() %>%
  setView(lat = 41.6, lng = -71.5, zoom = 8) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(data = voting_2020,
              stroke = TRUE,
              color = "purple",
              weight = 1,
              fillOpacity = 0) %>%
  addPolygons(data = voting_2016,
              stroke = TRUE,
              color = "orange",
              weight = 1,
              fillOpacity = 0)
```

