---
title: "Day Thirty-Five: Iteration"
subtitle: "SDS 192: Introduction to Data Science"
author: |
  Lindsay Poirier<br/>
  <span style = 'font-size: 70%;'>
  [Statistical & Data Sciences](http://www.smith.edu/sds), Smith College<br/>
  </span>
date: |
  Spring 2022<br/>
output: 
  html_document:
    toc: true
    toc_float: true
    theme: lumen
---

## Load Packages

```{r setup,	message = FALSE, warning = FALSE, include = FALSE}
library(tidyverse)
library(sf)
```

```{r}
import_state_data <- function(file_path, var_name){
  x <- st_read(file_path) %>%
    st_transform(4326) %>%
    mutate(TOTAL = rowSums(across(starts_with("G20")))) %>%
    mutate(PER_DEM = G20PREDBID / TOTAL * 100) %>%
    mutate(PER_DEM_LEAD = (G20PREDBID - G20PRERTRU) / TOTAL * 100) %>%
    st_set_geometry(NULL)
  
  assign(var_name, x, envir = .GlobalEnv)
  
  return(x)
}
```


```{r read-voting, include=FALSE}
states <- c("ma", "ct", "ri", "nh", "me", "vt")

var_names <- states %>% map_chr(~paste0("voting_2020_",.x))

file_paths <- states %>% map_chr(~paste0("Day35-Iteration/datasets/",.x, "_2020.shp"))

map2(file_paths, var_names, import_state_data)
```

# Write code to sum across all columns documenting total votes for a presidential candidate in `voting_2020_me`

```{r}
# Code here
```

# Write code to return the first 5 lines of `voting_2020_nh` in each county grouping

```{r}
# Code here
```

# Write a function to calculate the sum of votes for any candidate across all precincts

```{r}
sum_votes_any_cand<- function(df, var = G20PREDBID){
  df %>% summarize(Total = sum({{var}}))
}
```

```{r}
# Code here
```

# Write a function to create a histogram plotting the distribution of the percent dem lead across precincts in each New England state

```{r}
dem_lead_hist <- function(data, title, width_bins = 10){
 ggplot(data, aes(x = PER_DEM_LEAD)) +
    geom_histogram(binwidth = width_bins) +
   labs(title = paste("Distribution of Percent Dem Lead in", title, "2020"))
}
```

```{r}
# Code here
```

# Write a function that takes a town name in MA and determines how many votes were cast in that town

```{r}
votes_ma_towns <- function(town_name){

  x <- voting_2020_ma %>%
    filter(TOWN == town_name) %>%
    summarize(total_votes = sum(TOTAL))
  
  return(x)
}
```

```{r}
# Code here
```

# Write a function that takes a series of town names and returns a data frame of the precinct with the highest `PER_DEM_LEAD` in each town

```{r}
# Code here
```

```{r}
# Code here
```
