---
title: "Day Thirty-Four: Functions"
subtitle: "SDS 192: Introduction to Data Science"
author: |
  Lindsay Poirier<br/>
  <span style = 'font-size: 70%;'>
  [Statistical & Data Sciences](http://www.smith.edu/sds), Smith College<br/>
  </span>
date: |
  Spring 2022<br/>
output: 
  html_document:
    toc: true
    toc_float: true
    theme: lumen
---

## Load Packages

```{r setup,	message = FALSE, warning = FALSE, include = FALSE}
library(readr)
library(tidyverse)
library(sf)
```

```{r}
import_state_data <- function(file_path){
  x <- st_read(file_path) %>%
    st_transform(4326) %>%
    mutate(TOTAL = rowSums(across(starts_with("G20")))) %>%
    mutate(PER_DEM = G20PREDBID / TOTAL * 100) %>%
    mutate(PER_DEM_LEAD = (G20PREDBID - G20PRERTRU) / TOTAL * 100) %>%
    st_set_geometry(NULL)
  
  return(x)
}
```


```{r read-voting, include=FALSE}

voting_2020_ma <- import_state_data("Day34-Functions/datasets/ma_2020.shp")
voting_2020_ct <- import_state_data("Day34-Functions/datasets/ct_2020.shp")
voting_2020_ri <- import_state_data("Day34-Functions/datasets/ri_2020.shp")
voting_2020_vt <- import_state_data("Day34-Functions/datasets/vt_2020.shp")
voting_2020_nh <- import_state_data("Day34-Functions/datasets/nh_2020.shp")
voting_2020_me <- import_state_data("Day34-Functions/datasets/me_2020.shp")

```

# Write a function to calculate the sum of votes for Biden across all precincts

```{r}
sum_votes_any_cand<- function(data){
  
  x <- data %>%
    summarize(total_votes = sum({{num_votes}}))
  
  return(x)
}
```

```{r}
sum_votes_biden(voting_2020_ct)
sum_votes_biden(voting_2020_nh)
sum_votes_biden(voting_2020_ma)
sum_votes_biden(voting_2020_me)
sum_votes_biden(voting_2020_vt)
sum_votes_biden(voting_2020_ri)
```


# Adjust your function to take any candidate's column name as an argument. Set the default to votes for Biden. 

```{r}
sum_votes_any_cand(voting_2020_ct, G20PRERTRU)
sum_votes_any_cand(voting_2020_nh, G20PRERTRU)
sum_votes_any_cand(voting_2020_ma, G20PRERTRU)
sum_votes_any_cand(voting_2020_me, G20PRERTRU)
sum_votes_any_cand(voting_2020_vt, G20PRERTRU)
sum_votes_biden(voting_2020_ri, G20PRERTRU)
```

# Write a function to create a histogram plotting the distribution of the percent dem lead across precincts in each New England state

```{r}
dem_lead_hist <- function(data, width_bins = 10, title){
 x <- ggplot(data, aes(x = PER_DEM_LEAD)) +
    geom_histogram(binwidth = width_bins) +
   labs(title = title)
 
 return(x)
}
```

```{r}
dem_lead_hist(voting_2020_ct, title = "Distribution...CT")
dem_lead_hist(voting_2020_ma, state = "MA")
dem_lead_hist(voting_2020_ri, state= "RI")
dem_lead_hist(voting_2020_nh, state = "NH")
dem_lead_hist(voting_2020_me, state = "ME")
dem_lead_hist(voting_2020_vt, 2, "VT")
```


# Write a function that takes a town name in MA and determines how many votes were cast in that town

```{r}
votes_ma_towns <- function(data = voting_2020_ma, town_name){

  x <- data %>%
    filter(TOWN == town_name) %>%
    summarize(total_votes = sum(TOTAL))
  
  return(x)
}
```

```{r}
votes_ma_towns(town_name = "BLACKSTONE")
```

