---
title: "Day Nineteen: Importing Data"
author: Lindsay Poirier
subtitle: "SDS 192: Introduction to Data Science"
author: |
  Lindsay Poirier<br/>
  <span style = 'font-size: 70%;'>
  [Statistical & Data Sciences](http://www.smith.edu/sds), Smith College<br/>
  </span>
date: |
  Spring 2022<br/>
output: pdf_document
---

```{r}
library(readr)
library(stringr)
library(lubridate)
```

# Import `pioneer_valley_census_2013.csv`

Check your current working directory, and then find `pioneer_valley_census_2013.csv` in this project. Use the `read_lines()` function to check the first 5 lines of the this file. Be sure to reference the `?read_lines` help page to see how to limit the results to 5 lines.

```{r}
# Code here!
read_lines("Day7-GrammarGraphics/datasets/pioneer_valley_census_2013.csv", n_max = 5)
```

> Note how the column names do not appear on the first line. In what line do the column names appear? 

## Checkpoint 1: Import the data, skipping unnecessary rows so that the headers are the first line read in.

```{r}
# Code here!
pioneer_valley_census_2013 <-
  read_csv("Day7-GrammarGraphics/datasets/pioneer_valley_census_2013.csv", 
           skip = 2)
```

# Import NYC Bus Breakdown Data

Navigate to: https://data.cityofnewyork.us/Transportation/Bus-Breakdown-and-Delays/ez4e-fazm. Export the data as a CSV, create a `datasets` folder in the `Day19-Importing` folder in this project, and store the dataset in that folder.  

## Checkpoint 2: Import NYC bus breakdown data below into a variable called `bus_breakdowns`

```{r}
# Code here!
bus_breakdowns <-
  read_csv("Day19-Importing/datasets/Bus_Breakdown_and_Delays.csv")
```

## Checkpoint 3: Convert all columns with dates/times in the dataset to a datetime format, and store the output in a new data frame called `bus_breakdown_dates_cleaned`. 

> You can do this for each date column, but as a bonus you might try using `mutate_at()`: https://dplyr.tidyverse.org/reference/mutate_all.html

```{r}
# Code here!

#Good
bus_breakdowns_cleaned <-
  bus_breakdowns %>%
  mutate(Occurred_On = mdy_hms(Occurred_On),
         Created_On = mdy_hms(Created_On),
         Informed_On = mdy_hms(Informed_On),
         Last_Updated_On = mdy_hms(Last_Updated_On))  %>%
  mutate(How_Long_Delayed = str_replace_all(How_Long_Delayed, "\\?", "")) %>%
  mutate(How_Long_Delayed = na_if(How_Long_Delayed, ""))

#Better
bus_breakdowns_cleaned <-
  bus_breakdowns %>%
  mutate_at(c("Occurred_On", "Created_On", "Informed_On", "Last_Updated_On"), mdy_hms)

#Best
bus_breakdowns_cleaned <-
  bus_breakdowns %>%
  mutate_at(vars(matches("d_On")), mdy_hms)
```

There are a few rows in the `How_Long_Delayed` column with question marks `???` (some have more question marks than others). Can you identify those columns with data wrangling verbs? (Hint: You will need a string function here, and you may wish to refer to the Subsetting slides from last week.)

> Note that question marks serve a special purpose in `R` (help pages), and here we don't want to use the question mark for that special purpose. To treat the question mark as a regular character, we need to **escape** its special purpose. We can do that by placing two blackslashes in front of our reference to it, like this: '\\?'. The backslashes tell `R` to ignore its special purpose and treat it like any normal character. Other characters that share this feature include dollar signs, periods, backticks, etc. 

```{r}
bus_breakdowns_cleaned %>%
  filter(str_detect(How_Long_Delayed, "\\?")) %>%
  select(How_Long_Delayed)
```

## Checkpoint 4: Using a pipe to string a new function onto your date cleaning function above, replace all question marks in these rows with an empty string. Rerun the code you wrote to check for question marks above to ensure the question marks have been removed. 

> By converting all question marks to an empty string in `How_Long_Delayed`, certain cells in your data frame now have nothing in them but are not yet designated as `NA`. They are basically just character values with no text (or ""). Run the code below to see what I mean.

```{r}
# Code here!
bus_breakdowns_cleaned %>%
  filter(How_Long_Delayed == "") %>%
  select(How_Long_Delayed) %>%
  head()
```

## Checkpoint 5: Using a pipe to string a new function onto your date cleaning function above, convert empty text cells in `How_Long_Delayed` to `NA`s. Rerun the code I provided to check for empty strings above to ensure the question marks have been removed.

# Challenge

What steps might you take to clean up the `How_Long_Delayed` column so that all of the values are in a standard format? Note that this is a CHALLENGING task. I stared at the values in this column a while and only have mediocre solutions. If you want to practice, I recommend taking things one step at a time. 

