---
title: "Day Seven: Subsetting and Aggregating Data"
author: Lindsay Poirier
subtitle: "SDS 192: Introduction to Data Science"
author: |
  Lindsay Poirier<br/>
  <span style = 'font-size: 70%;'>
  [Statistical & Data Sciences](http://www.smith.edu/sds), Smith College<br/>
  </span>
date: |
  Spring 2022<br/>
output: pdf_document
---

```{r}
library(tidyverse)
library(lubridate)
hospitals <- read.csv("https://raw.githubusercontent.com/SDS-192-Intro/SDS-192-public-website/main/slides/datasets/Hospital_Cost_Report_2018.csv") %>%
  mutate(
    Fiscal.Year.Begin.Date = ymd(Fiscal.Year.Begin.Date),
    Fiscal.Year.End.Date = ymd(Fiscal.Year.End.Date))
```

# Which city in the US has the most hospitals represented in this dataset?

> Hint: Remember that different states might have cities with the same name! 

```{r}
hospitals %>%
  group_by(City, State.Code) %>%
  summarize(Total_Hospitals = n()) %>%
  arrange(desc(Total_Hospitals))
```

# In California, are there more rural or urban hospitals? Do rural or urban hospitals have more beds? Be sure to check for the number of hospitals with bed numbers missing. 

```{r}
hospitals %>%
  filter(State.Code == "CA") %>%
  group_by(Rural.Versus.Urban) %>%
  summarize(
    Total_Hospitals = n(),
    Total_Beds = sum(Number.of.Beds, na.rm = TRUE),
    Total_Beds_Missing = sum(is.na(Number.of.Beds))) 
```


# Critical Access Hospitals (CCN.Facility Type: CAH) are expected to have 25 beds or fewer. In each state, which critical access hospital has the most number of beds, and how many beds does it have? Return a data frame listing the hospital's name, state, and number of beds, and order the data frame from most beds to least.

```{r}
hospitals %>%
  filter(CCN.Facility.Type == "CAH") %>%
  group_by(State.Code) %>%
  filter(Number.of.Beds == max(Number.of.Beds, na.rm = TRUE)) %>%
  ungroup() %>%
  select(Hospital.Name, State.Code, Number.of.Beds) %>%
  arrange(desc(Number.of.Beds))
```
